---
title: "Abalones"
author: "Melinda Higgins"
date: "January 20, 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
# set up knitr options for all code chunks
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error = FALSE)

# load packages needed for code chunks
library(readr)
library(tibble)
library(dplyr)
library(ggplot2)

# import abalone.csv dataset
# use read_csv()
# function from readr to import data
abalone <- read_csv(file = "abalone.csv")
```

## A Glimpse of the Abalone Dataset

Use the `glimpse()` function from the `tibble` package to take a peak at the **abalone** dataset.

```{r dataglimpse}
# get a glimpse of the data
# use glimpse() function from tibble package
glimpse(x = abalone)
```

## Summary of abalone dimensions

1. Select three variables from the **abalone** dataset: _length_, _diameter_, and _height_.
2. Run the `summary()` function to get some basic descriptive statistics.

```{r summary}
# select a few variables
# and get summary stats
abalone %>% 
  select(length, diameter, height) %>% 
  summary()
```

## Plot of number of rings by shell weight

Make a scatterplot of the number of rings (on y-axis) by shell weight (on x-axis) and put plots in 3 separate panels by sex (i.e., male, female and infant) and add a smoothed fit line to each plot.

Note: The number of rings range from `r min(abalone$rings)` to `r max(abalone$rings)`.

```{r plot}
# make plot of number of rings by shellWeight
ggplot(data = abalone, 
       aes(x = shellWeight, y = rings, color = sex)) +
  geom_point() +
  facet_wrap(vars(sex)) +
  geom_smooth()
```

## Regression of abalone rings by shell weight

We can save regression model results, then put the model coefficients from the `summary()` function into a nice table using the `kable()` function from the `knitr` package.

```{r regmodel}
model1 <- lm(formula = rings ~ shellWeight, 
             data = abalone)
sum_model1 <- summary(model1)
knitr::kable(sum_model1$coefficients)
```

